---
title: "Example-quality-measure-analysis"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Example-quality-measure-analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(qmrel)
```


# Introduction
This vignette shows how to use the functions in qmrel to analzye quality measure performance.

First we'll simulate some data.

```{r}
n.providers = 100
n = rep(200, n.providers)
p = rbeta(n.providers, 0.5, 0.5) 
df <- simulateData(n = n, p = p)
```


# Calculate measure performance

```{r}
perf.results <- calcPerformance(df = df, y = "y", provider = "provider")
```

# Plot number of observations per provider

```{r}
plotN(perf.results$perf.results$n)
```

# Plot measure performance

```{r}
plotPerformance(perf.results$perf.results)
```


# Identify high- and low-performing entities


# Calculate risk-standardized performance


# Plot risk-standardized performance



# Identify high- and low-performing entities after risk-adjustment


# Calculate reliability

```{r}
rel.results <- calcReliability(df = df, y = "y", provider = "provider")
```



